---
title: "Joining Gapminder Data"
author: "Ojaswi Malik"
output: github_document
---

# Get the original data

```{r, echo = TRUE}
library(tidyverse)
library(gapminder)

gapminder
```

# Import the HIV data

```{r}
hiv_data <- read_csv("data/gapminder_hiv.csv")
```

# What is the relationship between HIV prevalence and life expectancy?

**Inner Join used** because the number of the data objects are greater in gapminder (X) than in hiv_data (Y) and using inner join will return all values in the X dataframe that have matching values in the Y dataframe. This way all the NA values for hivRate for countries that are present in gapminder and not in hiv_data can be removed and we can easily plot our graph.

```{r}

gapminder %>%
  inner_join(y = hiv_data) %>%
  ggplot(mapping = aes(x = hivRate, y = lifeExp)) +
  geom_smooth() +
  geom_point() +
  scale_x_log10() + 
  labs(
    title = "Relationship between HIV Prevalance and Life Expectancy",
    x = "HIV Prevalance Rate (Axis Logged) ",
    y = "Life Expectancy"
  )
  
```

# Which continents have the most observations with missing HIV data?

**Anti Join used** In order to narrow down those countries inside continents in gapminder (x) that do not have a corresponding HIV value in hiv_data, it is best to use anti_join. It is a filtering join which will return all rows in X that do not have a match in Y, which is exactly what we need.

```{r}
gapminder %>%
  anti_join(y = hiv_data) %>%
  count(continent) %>%
  ggplot(mapping = aes(x = reorder(continent, -n), y = n)) + 
  geom_col() + 
  labs(
    title = "Continents with missing HIV observations",
    x = "Continent",
    y = "Number of missing observations"
  )

```


## Session info

```{r, echo = TRUE}
devtools::session_info()
```


